\bf Title here

### 2020-07-07 paper/paper.Rmd
small tweaks here and there

The optimisation problem we're interested in is in the context of projection pursuit. Coined by @friedman1974projection, projection pursuit is a method that detects the interesting structure (i.e. clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions. Let $D_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $P = D \cdot A$, where $A_{p\times d}$ is the projection basis and $P_{n \times p}$ is the projected data. Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $A$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or indice function,  that can be used to measure the "interestingness" of the projection. A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection]. [Any literature on holes index? xxx]

@@

\begin{align}
I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search: $\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$. Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}$$

Here we use left subscript for the index of the interpolation at each iteration i,e, ${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration. The right subcript is reserved for the index of searching points in the optimisation algorithm.

There are three existing methods for optimisating PPI function and we review them below.

@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{j +1}_l$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$. If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops. If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted. $c$ is the halfing parameter meaning if $c > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search. The algorithm for a random search is summarised in Algorithm \ref{random-search}.

@@

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$. Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{log(l + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}$. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively. The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.

@@

\begin{equation}
\left[
\begin{array}{c|c|ccc|ccc}
\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
\hline
\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

where $n* = (J-2) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$ and $n^{**} = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$

@@

Below we will present several examples to diagnose different aspect of the tour optimisation. We will first provide examples on producing 1) static plots to explore the value of objective function and 2) animated plots to explore the searching space and then a more sophisticated example that combines both to understand how to optimise a noisy and complex index function. In both section, we will first provide a toy example that is easy to grasp and one or two examples that help us understand the optimisation algorithm and parameter choice better.

Follow the research question we raised earlier, the purpose of visual diagnostics is to understand

- Whether the procedure has provided an optimum and how the value of the objective function changes through the optimisation.

- How does the searching space look like, that is, geometrically, where are the points located in the space.

>>>>

The optimisation problem we're interested in is in the context of projection pursuit. Coined by @friedman1974projection, projection pursuit is a method that detects the interesting structure (i.e. clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions. Let $\mathbf{D}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{D} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times p}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection. A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection]. [Any literature on holes index? xxx]

@@

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search: $\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$. Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}$$

Where $\mathbf{A}^0$ is the starting basis and $\mathbf{A}^J$ is the final basis that has the largest index value in the searching space. Here we use left subscript to denote the index of the interpolation bases in each iteration i.e. ${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration. The right subcript is reserved for the index of searching points in the optimisation algorithm.

There are three existing methods for optimisating PPI function and we review them below. @posse1995projection proposed a stochastic direct search method, a random search algorithm. Given a current basis $\mathbf{A}^{j}$, a candidate basis $\mathbf{A}^{j +1}_l$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}^{j}$ by $$\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$$. If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $\mathbf{A}^{j + 1}$ along with other metadata and the current iteration stops. If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted. $c$ is the halfing parameter and when $c > 30$, the nieghbourhood parameter $\alpha$ in the next iteration will be reduced by half. The algorithm for a random search is summarised in Algorithm \ref{random-search}.

@@

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$. Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{log(l + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}$. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively. The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.

@@

\begin{equation}
\left[
\begin{array}{c|c|ccc|ccc}
\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
\hline
\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
\hline 
\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
\hline
{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
\hline
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n-1 &o_{1, n-1}& \cdots & o_{p, n-1}\\
\mathbf{A}^J & I^J & J & l_H + 1 & n & o_{1, n}& \cdots & o_{p, n}\\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

where $n = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$

@@

Below we will present several examples to diagnose different aspects of the projection pursuit optimisation. We will present:

1) static plots to explore the index value,

2) animated plots to explore the projection basis and,

3) a self-contained example to understand how to optimise a noisy and complex index function.

In the first two sections, we will first provide a toy example that is easy to grasp and then more examples that can help us to understand the algorithm and parameter choice. Remeber the research question we raised earlier, the purpose of visual diagnostics is to understand:

- Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm?

- How does the searching space look like, that is, geometrically, where are the projection bases located in the space?

### 2020-07-07 paper/paper.tex
small tweaks here and there

The optimisation problem we're interested in is in the context of
projection pursuit. Coined by \citet{friedman1974projection}, projection
pursuit is a method that detects the interesting structure
(i.e.~clustering, outliers and skewness) of multivariate data via
projecting it in lower dimensions. Let \(D_{n \times p}\) be a data
matrix, an n-d projection can be seen as a linear transformation:
\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(P = D \cdot A\),
where \(A_{p\times d}\) is the projection basis and \(P_{n \times p}\)
is the projected data. Define
\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
that maps the projection basis \(A\) onto a scalar value \(I\), this
function is commonly known as the projection pursuit index (PPI)
function, or indice function, that can be used to measure the
``interestingness'' of the projection. A number of indice functions have
been proposed in the literature and this includes Legendre index
\citep{friedman1974projection}, hermite index
\citep{hall1989polynomial}, natural hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
index \citep{lee2010projection}. {[}Any literature on holes index?
xxx{]}

@@

\begin{align}
I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search:
\(\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
Guided tour interpolates geodesically between each basis found by
projection pursuit \(A^j\) to form a tour path:
\[\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}\]

Here we use left subscript for the index of the interpolation at each
iteration i,e, \({}_{2}\mathbf{A}^1\) means the second interpolation
basis in the first iteration. The right subcript is reserved for the
index of searching points in the optimisation algorithm.

There are three existing methods for optimisating PPI function and we
review them below.

\citet{posse1995projection} used a stochastic direct search method, a
random search algorithm to sample a candidate bases \(A^{j +1}_l\) in
the neighbourhood of the current basis \(A^j\) defined by the radius of
the p-dimensional sphere, \(\alpha\). If the candidate basis has a
higher index value than the current basis, it is outputed as the target
basis \(A^{j + 1}\) along with other metadata and the algorithm stops.
If no basis is found to have higher index value after the maximum number
of tries \(l_{\max}\), the algorithm stops with nothing outputted. \(c\)
is the halfing parameter meaning if \(c > 30\), the nieghbourhood
parameter \(\alpha\) will be reduced by half in the next iteration of
the random search. The algorithm for a random search is summarised in
Algorithm \ref{random-search}.

@@

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
is defined as \(T(l) = \frac{T_0}{log(l + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(A_{rand}\), a candidate
basis is constructed as
\(A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
of the candidate basis is larger than the one of the current basis, the
candidate basis becomes the target basis. If it is smaller, the
candidate is accepted with probability
\(\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
where \(I^j\) and \(I^{j+1}_l\) are the index value of the current and
candidate basis respectively. The algorithm can be summarised as in
Algorithm \ref{simulated_annealing}.

@@

\begin{equation}
\left[
\begin{array}{c|c|ccc|ccc}
\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
\hline
\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

where \(n* = (J-2) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\) and
\(n^{**} = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\)

@@

Below we will present several examples to diagnose different aspect of
the tour optimisation. We will first provide examples on producing 1)
static plots to explore the value of objective function and 2) animated
plots to explore the searching space and then a more sophisticated
example that combines both to understand how to optimise a noisy and
complex index function. In both section, we will first provide a toy
example that is easy to grasp and one or two examples that help us
understand the optimisation algorithm and parameter choice better.

Follow the research question we raised earlier, the purpose of visual
diagnostics is to understand

\begin{itemize}
\item
  Whether the procedure has provided an optimum and how the value of the
  objective function changes through the optimisation.
\item
  How does the searching space look like, that is, geometrically, where
  are the points located in the space.
\end{itemize}

>>>>

The optimisation problem we're interested in is in the context of
projection pursuit. Coined by \citet{friedman1974projection}, projection
pursuit is a method that detects the interesting structure
(i.e.~clustering, outliers and skewness) of multivariate data via
projecting it in lower dimensions. Let \(\mathbf{D}_{n \times p}\) be a
data matrix, an n-d projection can be seen as a linear transformation:
\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
\(\mathbf{P} = \mathbf{D} \cdot \mathbf{A}\), where
\(\mathbf{P}_{n \times p}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis Define
\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
that maps the projection basis \(\mathbf{A}\) onto a scalar value \(I\),
this function is commonly known as the projection pursuit index (PPI)
function, or index function and can be used to measure the
``interestingness'' of the projection. A number of indice functions have
been proposed in the literature and this includes Legendre index
\citep{friedman1974projection}, hermite index
\citep{hall1989polynomial}, natural hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
index \citep{lee2010projection}. {[}Any literature on holes index?
xxx{]}

@@

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search:
\(\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
Guided tour interpolates geodesically between each basis found by
projection pursuit \(A^j\) to form a tour path:
\[\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}\]

Where \(\mathbf{A}^0\) is the starting basis and \(\mathbf{A}^J\) is the
final basis that has the largest index value in the searching space.
Here we use left subscript to denote the index of the interpolation
bases in each iteration i.e. \({}_{2}\mathbf{A}^1\) means the second
interpolation basis in the first iteration. The right subcript is
reserved for the index of searching points in the optimisation
algorithm.

There are three existing methods for optimisating PPI function and we
review them below. \citet{posse1995projection} proposed a stochastic
direct search method, a random search algorithm. Given a current basis
\(\mathbf{A}^{j}\), a candidate basis \(\mathbf{A}^{j +1}_l\) is sampled
in the neighbourhood defined by the radius of the p-dimensional sphere,
\(\alpha\) of the current basis \(\mathbf{A}^{j}\) by
\[\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}\]. If
the candidate basis has a higher index value than the current basis, it
is outputed as the target basis \(\mathbf{A}^{j + 1}\) along with other
metadata and the current iteration stops. If no basis is found to have
higher index value after the maximum number of tries \(l_{\max}\), the
algorithm stops with nothing outputted. \(c\) is the halfing parameter
and when \(c > 30\), the nieghbourhood parameter \(\alpha\) in the next
iteration will be reduced by half. The algorithm for a random search is
summarised in Algorithm \ref{random-search}.

@@

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
is defined as \(T(l) = \frac{T_0}{log(l + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(A_{rand}\), a candidate
basis is constructed as
\(A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
of the candidate basis is larger than the one of the current basis, the
candidate basis becomes the target basis. If it is smaller, the
candidate is accepted with probability
\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
where \(I^j\) and \(I^{j+1}_l\) are the index value of the current and
candidate basis respectively. The algorithm can be summarised as in
Algorithm \ref{simulated_annealing}.

@@

\begin{equation}
\left[
\begin{array}{c|c|ccc|ccc}
\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
\hline
\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
\hline 
\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
\hline
{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
\hline
\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n-1 &o_{1, n-1}& \cdots & o_{p, n-1}\\
\mathbf{A}^J & I^J & J & l_H + 1 & n & o_{1, n}& \cdots & o_{p, n}\\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

where \(n = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\)

@@

Below we will present several examples to diagnose different aspects of
the projection pursuit optimisation. We will present:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  static plots to explore the index value,
\item
  animated plots to explore the projection basis and,
\item
  a self-contained example to understand how to optimise a noisy and
  complex index function.
\end{enumerate}

In the first two sections, we will first provide a toy example that is
easy to grasp and then more examples that can help us to understand the
algorithm and parameter choice. Remeber the research question we raised
earlier, the purpose of visual diagnostics is to understand:

\begin{itemize}
\item
  Whether the algorithm has successfully found the maximum and how the
  index value changes throughout the algorithm?
\item
  How does the searching space look like, that is, geometrically, where
  are the projection bases located in the space?
\end{itemize}
